Who does big team science?

### 2024-01-16 manuscript_scopus.Rmd
updating
changes to resubmit

@@

According to the Oxford English dictionary, collaboration is two or more
people working together to achieve a certain goal [@oed2016].
Collaboration in scientific endeavors involves multiple researchers at
(potentially) multiple institutions to communicate and work together to
advance knowledge in their chosen field. Collaboration can manifest
uniquely in each project dependent on the skill sets, hypotheses, and
perspectives of collaborators. While collaboration is not new in
science, the current interest of "big team science" is increasing
[@coles2022; @forscher2022a; @stewart2017]. Big team science projects
and/or organizations utilize and run on large-scale collaboration to
ensure that diverse populations and ideas are brought into research
projects, which in turn allows for more reliability and generalizability
in the results and method of the study.

BTS appears to be expanding as a result of two sources: 1) increasing
globalization and technology that allows for real-time interdisciplinary
research, and 2) expanding interest in reproducibility, replication, and
generalizability [@maxwell2015; @nelson2018; @zwaan2018]. Technological
advances have provided easier ways to collaborate with people who are
from other universities and countries through document sharing platforms
(e.g., Google, GitHub, and the Open Science Framework), video chatting
platforms (e.g., Zoom, Microsoft Teams), and messaging and project
management platforms (e.g., Slack, Trello, when2meet, etc.). The
credibility movement seems to suggest that by having both collaborations
that span across the globe and subfields of research areas, age groups,
and education levels should help to drive science in the path of better
materials, reliability, generalizability and more robust sample sizes
(when necessary) in a study [@auspurg2021; @nosek2014method;
@lebel2018].

The credibility movement was originally defined by a focus on large
scale replications used in collaborative environments [@vazire2022].
Generally, the movement appears to have been driven by early career
researchers (i.e., those who are within five years of their first
appointment) [@maizey2019]; however, there are no large meta-scientific
investigations on this specific topic to date. Potentially, the lack of
investigation is tied to the newness of the large-scale research in many
fields, as it is only in recent years that publications like the Open
Science Collaboration [@opensciencecollaboration2015], Many Labs
Collaborations [@buttrick2020; @ebersole2016; @ebersole2020; @klein2018;
@klein2022; @mathur2020; @skorb2020] or the first papers from the
Psychological Science Accelerator [@bago2022; @buchanan2023;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@moshontz2018; @wang2021]. Generally, the researcher incentive for
replication and/or involvement in big-team projects was low for three reasons. First, journals often prioritize
"novel" or new results which led to rejection of replication manuscripts
and publication bias [@franco2014; @hubbard1997; @nosek2012]. Second,
the "failure" to replicate was often placed on the replication team as
"bad science" rather than a careful consideration of publication biases
and (potential) questionable research practices [@klein2022;
@maxwell2015]. Last, why should someone want to spend time and resources
on an answer we already "know" [@isager2021; @isager2021a]?

@@

Beyond the replication movement, the credibility movement has mirrored calls for diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research [@henrich2010; @newson2021;
@rad2018] by improving representation in research samples. Like the
large-scale studies in Physics [@aphilos2021; @castelnovo2018] and
Biology [@collins2003], the Social Sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@bago2022;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@vanbavel2022; @wang2021]. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly *who* is involved with,
and benefits from, the BTS and credibility movement. Publications on BTS
generally explore challenges, lessons learned, and the need for BTS
[@forscher2022a; @coles2022].

Therefore, the goal of this manuscript is to examine both the *publications* and *people* involved in BTS projects. We present descriptive information about the publication sources and types of articles that we classified as BTS projects to demonstrate what areas of research show large-scale research. Next, we examine the individuals involved in those projects for descriptive and predictive purposes. To describe the people involved in BTS projects, we used education, types of publications (i.e., articles, preprints, books, etc.) from BTS individuals, and publication metrics.

For predictive statistics, we explored the change in diversity of authors over time. We see an increasing interest and number of publications in BTS but we do not yet
know if this uptick in large-scale projects has diversified the *people*
involved in BTS. While a few publications have noted that BTS appears to
be early career researchers [@maizey2019], no one has systematically
investigated this perception. Further, it is unclear if the focus of
de-WEIRDing science has only focused on the representation of the
research participants or if it has also improved the representation of
researchers outside of North America and Europe. Last, who runs these
BTS projects? Do we see an increase in diversity for the authors who
generally receive the most credit for these projects (i.e., first
several author(s) and last author)? As hiring and promoting practices
often place a heavy weight on publications and especially "influential"
publications, it becomes necessary to critically examine the
representation present in authorship in BTS projects.

@@

This manuscript was preregistered with the same conceptual ideas using
Google Scholar and ORC-ID databases (MASKED FOR REVIEW^[Unfortunately, the way we created these pre-registrations did not allow for complete anonymization required by the journal]) but then
was updated with access to the Scopus database for a broader picture of
BTS projects (MASKED FOR REVIEW). All materials and code can be
found on our OSF page: MASKED FOR REVIEW or corresponding GitHub
archive: MASKED FOR REVIEW.

@@

### Number of articles

@@

### Number of journals

@@

In this investigation, we explored the publication rates, areas, and
researchers involved in big team science publications. Over a
half-million articles were published in nearly 15,000 journals since
1970 that qualified as big team science articles (at least ten authors
and ten different affiliations). The areas of publication were aligned
to cancer and genetics research in medicine and oncology for Health and
Life Sciences, physics and chemistry for the Physical Sciences, and
psychology for the Social Sciences All areas of research show an
explosion in growth in the number of publications and the number of
authors included on manuscripts, replicating previous investigations
into this topic area [@sinatra2015; @wuchty2007; @hunter2008].

Our investigation expands into an exploration of the researchers who
publish in big teams focusing on diversity in seniority of authors and
geopolitical affiliation. The number of earlier career scholars is
increasing in publications across the years, indicating that big teams
may be more accessible to different types of individuals, not just
older, more established researchers. This result is especially
interesting given the publish-or-perish model still present in most
institutions, as it may seem that large-scale projects could be a risky
choice for non-permanent researchers. In the authors' experience, big
team projects are often quite slow to publication, incentives may be low
for non corresponding authors if institutions do not value papers
without lead authorship, and there is no guarantee for publication even
with a large group. However, with a large team the distribution of work
could imply less effort on individual non-leading members, and research
has shown that larger-team publications do receive more citations and
appear to have higher impact [@larivière2015].

The results for the number of publications by big team researchers
mimics the findings from career length, with a smaller effect size. In
general, it appears that there is a decrease in the average number of
publications a researcher has when publishing on a big team science
paper over time. This result is likely attributable to the number of
early career scholars joining projects, but also may support increased
accessibility for individuals to be involved in this type of research.
Globalization, the internet, and the focus on interdisciplinary research
are potentially driving forces behind our results, but, hopefully, the
results also point to a decline in scientific gatekeeping [@lu2007;
@siler2015].

@@

The number of geopolitical entities for researcher affiliation is
increasing over time, showing the results of globalization and the
ability to connect across time zones and cultures [@xie2014]. While our
definition of big team science required at least ten different
institutional affiliations, we did not filter papers by geopolitical
region, and thus, a manuscript could rely solely on institutions within
a single country. The Physical Sciences did not show an increase in
diversity of regions represented, however, it could be argued that the
development of large research centers like CERN forced earlier diversity
than other sciences (i.e., because CERN specifically recruited
scientists from sponsoring nations). The Life, Health, and Social
Sciences saw an increase in the number of regions represented with the
highest increase in the Social Sciences. This result likely corresponds
with an increased interest in big team science publications in
psychology [@coles2022; @forscher2022a], and the desire to diversify the
populations represented in psychological research [@henrich2010;
@newson2021].

While publications on the whole are diversifying, we did not yet find
equality in the representation for first/corresponding author spots
versus all other authors. In general, first authors appear to be less
diverse, representing more European and North American authors, while
other authors include more Asian and African authors. These effect sizes
were often small, but the inequality still persists across years, even
if they are slowly decreasing. Diverse teams are more likely to have
papers with stronger "impact" [@freeman2015; @hinnant2012; @jones2008;
@yang2022] with higher citation metrics for more diverse author lists.
The introduction of contributorship models (e.g., CRediT[@allen2019])
will hopefully continue to push these effects down, as they highlight
each individual's contribution to a manuscript.

The limitations for this research are tied to the availability and
curation of the Scopus dataset. While the number of articles analyzed
for this investigation was large, the criteria for inclusion requires
the correct entry of author affiliations, the correct author linkages
for career length and publication rates, and the correctly marked
geopolitical entity. We had planned to analyze educational levels to
determine if the number of student coauthors (i.e., non-terminal degree)
had increased over time; however, this data was mostly blank within the
Scopus archive. Scopus is a carefully curated dataset, but these
limitations must be kept in mind when interpreting the results.
Publication language diversity was not investigated, and a previous
study indicates that the majority of publications in big databases are
in English [@albarillo2014]. Certainly, publications in non-English
languages would improve the statistics on diversity in scientific
publishing - but the English language barrier likely exists regardless
of inclusion in databases [@meneghini2007; @ramírez-castañeda2020].

Big teams have the ability to provide high-impact, important research
within scientific publishing, and this report suggests a promising trend
of increasing numbers of publications that include earlier career and
more diverse scholars. These partnerships introduce new challenges to
collaboration from interpersonal conflict, infrastructure, incentives,
to international political situations [@forscher2022a]. Directed studies
into ways to navigate these situations would be beneficial for policy
makers at institutions, as well as lead teams who organize and complete
these projects. The implications for retention and promotion processes
across a broad span of regions should be explored to improve diversity
with the understanding of the differential impact of incentives for
participating in big team studies.

@@

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

>>>>

  # General Disclosures

  Conflicts of interest: Both authors declare no conflicts of interest.

  Funding: No funding was obtained for this project.

  Artificial intelligence: No AI resources were used for this manuscript.

  Ethics: No ethics review was necessary for this project.

  Computational reproducibility: All materials and code can be found on our OSF page: MASKED FOR REVIEW or corresponding GitHub archive: MASKED FOR REVIEW. Elsevier has agreed to provide access to determine reproducibility of the code for accessing and summarizing articles, and the reproducible manuscript has been provided for review.

  Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (MASKED FOR REVIEW^[Unfortunately, the way we created these pre-registrations did not allow for complete anonymization required by the journal, this work has been uploaded a deidentified zip file]) but then was updated with access to the Scopus database for a broader picture of BTS projects (MASKED FOR REVIEW).

  Materials: MASKED FOR REVIEW (see uploaded zip file).

  Data: MASKED FOR REVIEW (see uploaded zip file).

  Analysis scripts: MASKED FOR REVIEW (see uploaded zip file).

@@

Collaboration in scientific endeavors involves multiple researchers at
(potentially) multiple institutions to communicate and work together to
advance knowledge in their chosen field. Collaboration can manifest
uniquely in each project dependent on the skill sets, hypotheses, and
perspectives of collaborators. While collaboration is not new in
science, the interest in "big team science" is increasing
[@coles2022; @forscher2022a; @stewart2017]. Big team science (BTS) projects
and/or organizations utilize and run large-scale collaborations to
ensure that diverse populations and ideas manifest in research
projects, which in turn allows for more reliability and generalizability
in the results and methods of the study.

BTS appears to be expanding because of two sources: 1) increasing
globalization and technology that allows for real-time interdisciplinary
research, and 2) expanding interest in reproducibility, replication, and
generalizability [i.e., the credibility movement, 
@maxwell2015; @nelson2018; @zwaan2018; @vazire2022]. Technological
advances have provided easier ways to collaborate with people who are
from other universities and countries through document sharing platforms
(e.g., Google, GitHub, and the Open Science Framework), video chatting
platforms (e.g., Zoom, Microsoft Teams), and messaging and project
management platforms (e.g., Slack, Trello, when2meet, etc.). The
credibility movement seems to suggest that by having both collaborations
that span across the globe and subfields of research areas, age groups,
and education levels should help to drive science in the path of better
materials, reliability, generalizability, and more robust sample sizes
(when necessary) in a study [@auspurg2021; @nosek2014method;
@lebel2018].

Generally, the credibility movement appears to have been driven by early career
researchers [i.e., those who are within five years of their first
appointment, @maizey2019]; however, there are no large meta-scientific
investigations on this specific topic to date. The newness of large-scale research in many fields could be the culprit for the lack of investigation into this area. For example, psychology has had an increase in BTS publications like the Open
Science Collaboration [@opensciencecollaboration2015], Many Labs
Collaborations [@buttrick2020; @ebersole2016; @ebersole2020; @klein2018;
@klein2022; @mathur2020; @skorb2020] or the first papers from the
Psychological Science Accelerator [@bago2022; @buchanan2023;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@moshontz2018; @wang2021]. Generally, the researcher incentive for
replication and/or involvement in big-team projects was low for three reasons. First, journals often prioritize
"novel" or new results which led to rejection of replication manuscripts
and publication bias [@franco2014; @hubbard1997; @nosek2012]. Second,
the "failure" to replicate was often placed on the replication team as
"bad science" rather than a careful consideration of publication biases
and (potential) questionable research practices [@klein2022;
@maxwell2015]. Last, why should someone want to spend time and resources
on an answer we already "know" [@isager2021; @isager2021a]?

@@

Beyond replication concerns, the credibility movement has mirrored calls for diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research [@henrich2010; @newson2021;
@rad2018] by improving representation in research samples. Like the
large-scale studies in Physics [@aphilos2021; @castelnovo2018] and
Biology [@collins2003], the Social Sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@bago2022;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@vanbavel2022; @wang2021]. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly *who* is involved with,
and benefits from, the BTS and credibility movement. Publications on BTS
generally explore challenges, lessons learned, and the need for BTS
[@forscher2022a; @coles2022].

Therefore, the goal of this manuscript is to examine both the *publications* and *people* involved in BTS projects. We present descriptive information about the publication sources and types of articles to demonstrate large-scale research. Next, we examine the individuals involved in BTS for descriptive and predictive purposes. To describe the people involved in BTS projects, we planned to use education, types of publications from BTS individuals, and publication metrics. For predictive statistics, we explored the change in diversity of authors over time. It is unclear if the focus of de-WEIRDing science has only focused on the representation of the
research participants or if it has also improved the representation of
researchers outside of North America and Europe. Last, we examined for a change in diversity within first author(s) and the last author across time. As hiring and promoting practices often place a heavy weight on publications and especially “influential” publications, it becomes necessary to critically examine the representation present in authorship in BTS projects.

@@

@@

### Number of Articles

@@

### Number of Journals

@@

In this investigation, we explored the publication rates, areas, and researchers involved in BTS publications. Over a half-million articles were published in nearly 15,000 journals since 1970 that qualified as BTS articles. The areas of publication were aligned to cancer and genetics research in medicine and oncology for Health and Life Sciences, physics and chemistry for the Physical Sciences, and psychology for the Social Sciences. All areas of research show growth in the number of publications and authors included on manuscripts, replicating previous investigations [@sinatra2015; @wuchty2007; @hunter2008].

Our investigation expands previous research by additionally focusing on diversity in seniority of authors and geopolitical affiliation. The number of earlier career scholars increased across years, indicating that big teams may be accessible to different types of individuals, not just older, established researchers. This result is especially interesting given the publish-or-perish model still present in most institutions, as it may seem that large-scale projects could be a risky choice for non-permanent researchers. BTS projects are often slow to publish, there is no guarantee for publication, and incentives are often lower for non-corresponding authors. However, with a large team, the distribution of work could imply less effort on individual non-leading members, and research has shown that larger-team publications do receive more citations and appear to have higher impact  [@larivière2015].

In general, it appears that there is a decrease in the average number of publications a researcher has when publishing in a BTS paper over time, mirroring career length results. This result is likely attributable to the number of early career scholars joining projects, but also may support increased accessibility for individuals to be involved. Globalization, the internet, and the focus on interdisciplinary research are potentially driving forces behind our results, but, hopefully, the results also point to a decline in scientific gatekeeping [@lu2007; @siler2015].

@@

The number of geopolitical entities for researcher affiliation is
increasing over time, showing the results of globalization and the
ability to connect across time zones and cultures [@xie2014]. While our
definition of BTS required at least ten different
institutional affiliations, we did not filter papers by geopolitical
region, and thus, a manuscript could rely solely on institutions within
a single country. The Physical Sciences did not show an increase in
diversity of regions represented, however, it could be argued that the
development of large research centers like CERN forced earlier diversity
than other sciences (i.e., because CERN specifically recruited
scientists from sponsoring nations). The Life, Health, and Social
Sciences saw an increase in the number of regions represented with the
highest increase in the Social Sciences. This result likely corresponds
with an increased interest in big team science publications in
psychology [@coles2022; @forscher2022a], and the desire to diversify the
populations represented in psychological research [@henrich2010;
@newson2021].

While publications overall are diversifying, we found differences in the representation for first/corresponding authors versus all other authors. In general, first authors appear to be less diverse, representing European and North American authors, while other authors include more Asian and African authors. These effect sizes were often small, but the inequality persists across years. Diverse teams are more likely to have papers with stronger “impact” [@freeman2015; @hinnant2012; @jones2008;
@yang2022] with higher citation metrics for more diverse author lists.
The introduction of contributorship models [e.g., CRediT, @allen2019]
will hopefully continue to push these effects down, as they highlight
each individual's contribution to a manuscript.

The limitations for this research are tied to the curation of the Scopus dataset: the correct author affiliations, the correct author publication information, and the correctly marked geopolitical entity. We had planned to analyze educational levels change over time; however, this data was mostly blank within the Scopus archive. Scopus is a carefully curated and large dataset, but these limitations must be kept in mind when interpreting the results. Publication language diversity was not investigated, and a previous study indicates that most publications in big databases are in English [@albarillo2014]. Certainly, publications in non-English
languages would improve the statistics on diversity in scientific
publishing - but the English language barrier likely exists regardless
of inclusion in databases [@meneghini2007; @ramírez-castañeda2020].

Big teams can provide high-impact, important research within scientific publishing, and this report suggests a promising trend of increasing publications that include earlier career and more diverse scholars. These partnerships introduce new challenges to collaboration from interpersonal conflict, infrastructure, incentives, to international political situations [@forscher2022a]. The implications for retention and promotion processes across a broad span of regions should be explored to improve diversity with the understanding of the differential impact of incentives for participating in big team studies.

@@

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}

```
::: {#refs custom-style="Bibliography"}
:::
